---
title: "DATA607_Assignment_3"
author: "Ying Fang Lee"
date: "9/22/2024"
output: html_document
---


## 1- Normalization

Created a csv file with sample table to be normalized [Sample Table.csv](https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_3/Normalize%20Table.csv) and uploaded to GitHub. This table has 7 columns that represents a students' grade for courses taught by different professors.  

####  Check if it is in first normal form
1) Each row is representing a different observation
2) No mix data types in each column
3) Does not store a repeating group of data items in one row
4) This table have a composite primary keys of Student_ID, Course_ID, and Teacher_Email
```{r}
library(tidyverse)

# Read the file from GitHub and save to a data frame
denormalized <- read.csv("https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_3/Normalize%20Table.csv")
print(denormalized)

```

####  Check if it is in second normal form
Second normal form, each non-key attribute must depend on the entire primary key
Student_Name depends on the Student_ID
Course_Name depends on the Course_ID
Teacher name each have unique email and they each teach a course
The bridge table has grade dependent on student, course, and teacher

```{r}
# Student_ID is the primary key for Student_Name
# Get the distinct Students from the table
Students <- denormalized %>% 
  select(Student_ID, Student_Name) %>% 
  distinct(Student_ID, .keep_all = TRUE)%>% 
  print()

# Course_ID is the primary key for Course_Name
# Get the distinct courses from the table
Courses <- denormalized %>% 
  select(Course_ID, Course_Name) %>% 
  distinct(Course_ID, .keep_all = TRUE)%>% 
  print()

# Teacher.email is the primary key for teacher and each teacher teaches a course
# Get the distinct teachers from the table
Teachers <- denormalized %>% 
  select(Teacher, Teacher.Email, Course_ID) %>% 
  distinct(Teacher.Email, .keep_all = TRUE) %>% 
  print()

# The de-normalized table after removing Student, Courses, and teacher with only the primary key of each table.
Bridge_Table <- denormalized %>% 
  select(Student_ID, Course_ID, Teacher.Email, Grade) %>% 
  print()
```

####  Check if it is in third normal form
Each non-key attribute must be dependent on the key, the whole key, and nothing but the key.

```{r}
# Looking at the Teachers table there is a relationship between Course_ID and Teachers.To represent that I created another bridge table called Teachers_Course and created an ID call Section_ID and created a primary key call Teacher_ID

# Modify Teachers table to add Teacher_ID
Teachers <- Teachers %>% 
  mutate(Teacher_ID = 1:n(), .before = Teacher) %>% 
  print()

# Create dataframe for Course_Section
Teacher_Course <- Teachers %>% 
  select(Course_ID, Teacher_ID) %>% 
  arrange(Course_ID) %>% 
  mutate(Section_ID = c(1,2,1)) %>% # manually creating the section_id for now
  print()

# converting the bridge table to Grades based on student, course, and section
Grades <- Bridge_Table %>% 
  left_join(Teachers) %>% 
  left_join(Teacher_Course) %>% 
  select(Student_ID, Course_ID, Section_ID, Grade) %>% 
  print()


```
#### Normalized tables
```{r}
print(Students)
print(Courses)
print(Teachers)
print(Teacher_Course)
print(Grades)

```

### 2 - Character Manipulation

Using the 173 majors listed in fivethirtyeight.com’s College [Majors dataset](https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/college-majors/majors-list.csv), provide code that identifies the majors that contain either "DATA" or "STATISTICS". 

There are three majors containing "DATA" or "STATISTICS" in the name.
```{r}
majors <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/college-majors/majors-list.csv")

str_view(majors$Major,"DATA|STATISTICS")
```

### 3. Describe, in words, what these expressions will match:

#### (.)\1\1 
This expression will match a string with three repeating characters. \
(.) Capturing group is of any character \
\1 same character as the capturing group repeat once \
\1 same character as the capturing group repeat the second time \
```{r}
x <- str_c("aaa bbb abc")
str_view(x, "(.)\\1\\1")
```
  
#### "(.)(.)\\2\\1"
This expression will match two characters followed by two characters in reversed order. \
(.) 1^st capturing group of any character \
(.) 2^nd capturing group of any character \
\\2 character from the second capturing group \
\\1 character from the first capturing group \
```{r}
x <- str_c("abba abbac aaaa")
str_view(x, "(.)(.)\\2\\1")
```
  
#### (..)\1
This expression will match string with any two characters repeated once. \ 
(..) capturing group of any two characters \
\1 repeat the characters from the capturing group once \

```{r}
x <- str_c("aaaa abababab cababc")
str_view(x, "(..)\\1")
```
  
#### "(.).\\1.\\1"
This expression match 5 characters where every other odd character is the same character \

(.) Capturing group of any character \
. any character \
\1 repeat of capturing group character \
. any character \
\1 repeat of capturing group character \
 
```{r}
x <- str_c("ababa abaca aaaaaa")
str_view(x, "(.).\\1.\\1")
```
  
#### "(.)(.)(.).*\\3\\2\\1"
This expression will match string with the three characters follow by any character and then the reverse of the three characters in the beginning \

(.) first capturing group of any character \
(.) second capturing group of any character \
(.) third capturing group of any character \
.* any number of characters \
\3 third capturing group of any character \
\2 second capturing group of any character \
\1 first capturing group of any character \
```{r}
x <- str_c("abcdefedcba efg gfe")
str_view(x,"(.)(.)(.).*\\3\\2\\1")

```
### 4. Construct regular expressions to match words that:
#### Start and end with the same character
```{r}
# "(.).*\\1"
x <- str_c("aba xyxdwn blob")
str_view(x, "(.).*\\1" )
```

#### Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
```{r}
# "(..).*\\1"
x <- ("church photography")
str_view(x, "(..).*\\1")
# If we want to tweak it to have letters starts and ends with the same letters
# "^(..).*\\1$"
x <- str_c("church")
str_view(x, "^(..).*\\1$")
```
  
#### Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
```{r}
# "(.).*\\1.*\\1"
x <- str_c("ababa cdceleven fififofum")
str_view(x, "(.).*\\1.*\\1")
```

