---
title: "Assignment_7"
author: "Ying Fang Lee"
date: "2024-10-20"
output: html_document
---

```{r}
library(tidyverse)
```
### Uploading the information as a csv file to Github then read it as a dataframe with R
I choose to tidyup the data before I create data files thinking it will be easier to read.
```{r}
# Read the csv file and clean the data into tidy form
raw <- "https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_7/data.csv"
products <- read.csv(raw)

# Replace . in column name to _
names(products) <- gsub("\\.","_", names(products))

# Separate the additional Column "X" for details from table 1
products_1 <- products %>%
  select(-c(X))
# Separate the Variation_Details column from X and rename X to Variation Details
products_2 <- products %>%
  select(-c(Variation_Details)) %>% 
  rename(Variation_Details = X)
# Combine table 1 and table two and separate the Variation_Details to two columns by :
products <- rbind(products_1, products_2) %>% 
  separate(Variation_Details, into = c("Variation_Type", "Variation_Detail"), sep = ":") %>% 
  arrange(Variation_ID) 
head(products)
```

## Create JSON, HTML, XML and Parquet with R
### Creating HTML file with the csv file as a table
```{r}
# Create HTML file with the content as a table

library(htmltools)
# Create HTML content
html_content <- tags$html(
  tags$body(
    tags$table(
      tags$thead(
        tags$tr(
          tags$th("Category"),
          tags$th("Item Name"),
          tags$th("Item ID"),
          tags$th("Brand"),
          tags$th("Price"),
          tags$th("Variation ID"),
          tags$th("Variation Type"),
          tags$th("Variation Details")
        )
      ),
      tags$tbody(
        lapply(1:nrow(products), function(i) {
          tags$tr(
            tags$td(products$Category[i]),
            tags$td(products$Item_Name[i]),
            tags$td(products$Item_ID[i]),
            tags$td(products$Brand[i]),
            tags$td(products$Price[i]),
            tags$td(products$Variation_ID[i]),
            tags$td(products$Variation_Type[i]),
            tags$td(products$Variation_Detail[i])
          )
        })
      )
    )
  )
)

# Save the HTML file
save_html(html_content, "product_data_report.html")
```

### Creating JSON file with the csv file
```{r}
# Create JSON file with the content from the data frame created from csv
library(jsonlite)

# Convert to JSON
json_data <- toJSON(products, pretty = TRUE)

# Save the JSON file
write(json_data, "product_data.json")
```

### Creating XML file with the csv file
```{r}
# Create XML file with the content from the data frame created from csv
library(XML)
# Create the XML document
xml_doc <- newXMLDoc()

# Create root node
root <- newXMLNode("Products", doc = xml_doc)

# Add product entries
for (i in 1:nrow(products)) {
  product_node <- newXMLNode("Product", parent = root)
  newXMLNode("Category", products$Category[i], parent = product_node)
  newXMLNode("Item_Name", products$Item_Name[i], parent = product_node)
  newXMLNode("Item_ID", products$Item_ID[i], parent = product_node)
  newXMLNode("Brand", products$Brand[i], parent = product_node)
  newXMLNode("Price", products$Price[i], parent = product_node)
  newXMLNode("Variation_ID", products$Variation_ID[i], parent = product_node)
  newXMLNode("Variation_Type", products$Variation_Type[i], parent = product_node)
  newXMLNode("Variation_Details", products$Variation_Detail[i], parent = product_node)
}
# Save the XML file
saveXML(xml_doc, file = "product_data.xml")
```

### Creating pqrquet file with the csv file

```{r}
# Create XML file with the content from the data frame created from csv
library(arrow)
# Write parquet file with dataframe
write_parquet(products,"product_data.parquet")
```
## Reading JSON, HTML, XML, parquet files

### Reading HTML file
```{r}
# Reading HTML file
library(rvest)
raw <- "https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_7/product_data_report.html"
html_file <- read_html(raw)

# Locates the table element and converts it to data frame
product_html <- html_file %>% 
  html_node("table") %>%
  html_table()

print(product_html)

```
### Reading JSON file
```{r}
# Reading JSON file
library(jsonlite)
raw <- "https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_7/product_data.json"

products_json <- fromJSON(raw)
print(products_json)


```
### Reading XML file
```{r}
# reading XML file
library(xml2)

raw <- "https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_7/product_data.xml"
xml_file <- read_xml(raw)

# Extract data from the XML
products <- xml_find_all(xml_file, ".//Product")

# Create a data frame to hold the extracted data
product_xml <- data.frame(
  Category = xml_text(xml_find_all(products, "Category")),
  Item_Name = xml_text(xml_find_all(products, "Item_Name")),
  Item_ID = xml_text(xml_find_all(products, "Item_ID")),
  Brand = xml_text(xml_find_all(products, "Brand")),
  Price = as.numeric(xml_text(xml_find_all(products, "Price"))),
  Variation_ID = xml_text(xml_find_all(products, "Variation_ID")),
  Variation_Type = xml_text(xml_find_all(products, "Variation_Type")),
  Variation_Details = xml_text(xml_find_all(products, "Variation_Details")),
  stringsAsFactors = FALSE
)

# View the data frame
print(product_xml)
```
### Reading parquet file
```{r}
# Reading parquet file
library(arrow)

# Unable to generate the raw file from Github
parquet_file <- "./product_data.parquet"
product_parquet <- read_parquet(parquet_file)

print(product_parquet)

```

## Conclusion JSON, HTML, XML, and parquet file
### HTML
HTML uses predefined tags to structure a web page. It has many more tags that describes webpage layout and not very suitable for data exchange. In general it will be harder to get data from HTML page since it contains other information that does not pertain to data.  

### JSON
JSON files is commonly used by APIs its file size is smaller compared to XML but larger than csv files. JSON files is loaded completely into memory which makes it slower to read. However, it is easy for human to read and great for unstructured data.  

### XML
XML is a markup language designed for data representation and storage. It contains user defined tags that makes it versatile for application. XML can be used to represent complex data structures. However, XML file is in general larger than JSON file since it can contain a lot of information.  

### Parquet
Parquet is a self-describing in that it includes metadata that includes the schema and structure of the file. It is one of the fastest file types to read, faster than JSON, and the file size is small. However, it is not easy to read like JSON files.  
