---
title: "Project_2"
author: "Ying Fang Lee"
date: "10/13/2024"
output: html_document
---

### My data

I choose the article [SO MUCH CANDY DATA, SERIOUSLY](https://www.scq.ubc.ca/so-much-candy-data-seriously/). This data is based on a survey conducted in 2017 on what people would like to get for Halloween. The orignal survey looks like [this] (https://www.scq.ubc.ca/wp-content/uploads/2017/10/candyhierarchysurvey2017.pdf).  The data recorded was very wide with some manual entries for country and state.There are also a lot of missing data where the question was not answered completely. 

I am interested to find out the kind of candy people wish to get on Halloween in 2017 and create a ranking for it. 

### Issues
- Biggest problem is data corruption caused by human error. Spelling mistakes, answer is not uniform, and NULL answers.
- Counties and States not spelled correctly will be part of the margin of error and be omitted from analysis.
- Preference is not numerical
- Data set where answers are incomplete.
- Columns contain unidentified characters.


```{r}
library(tidyverse)
url <-"https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Project_2/candyhierarchy2017.csv"
file<- read.csv(url, fileEncoding = "UTF-8")
# Taking care of empty spaces and set them to NA
file[file ==""] <- NA
# Changing the ranking for candy to number for analysis later
file[file=="MEH"] <- 2
file[file=="JOY"] <- 3
file[file=="DESPAIR"] <- 1

glimpse(file)
```

### Cleaning Up States column

-   Match pattern with State Names spelled out to the two letter State abbreviations -   Parse the cell to extract the states
-   Leave all the other states from other countries alone
```{r}
US_States <- file$Q5..STATE..PROVINCE..COUNTY..ETC

# State Patterns

sname <- c("District of Columbia" ,  "Alabama"  , "Alaska"  , "American Samoa" ,   "Arizona" ,  "Arkansas" ,  "California" , "Colorado" , "Connecticut" , "Delaware" , "Florida" , "Georgia" , "Guam" , "Hawaii"  ,  "Idaho"  ,  "Illinois"  ,  "Indiana"  ,  "Iowa" , "Kansas" , "Kentucky" , "Louisiana" , "Maine" , "Maryland" , "Massachusetts" , "Michigan" , "Minnesota" , "Mississippi" , "Missouri" , "Montana" , "Nebraska" , "Nevada" , "New Hampshire" , "New Jersey" , "New Mexico" , "New York" , "North Carolina" , "North Dakota" , "Ohio" , "Oklahoma" , "Oregon" , "Pennsylvania" , "Puerto Rico" , "Rhode Island" , "South Carolina" , "South Dakota" , "Tennessee" , "Texas" , "Utah" , "Vermont" , "Virginia" , "Washington" , "West Virginia" , "Wisconsin" , "Wyoming")
states2 <- c("DC" , "AL" , "AK" , "AS" , "AZ" , "AR" , "CA" , "CO" , "CT" , "DE" , "FL" , "GA" , "GU" , "HI" , "ID" , "IL" , "IN" , "IA" , "KS" , "KY" , "LA" , "ME" , "MD" , "MA" , "MI" , "MN" , "MS" , "MO" , "MT" , "NE" , "NV" , "NH" , "NJ" , "NM" , "NY" , "NC" , "ND" , "OH" , "OK" , "OR" , "PA" , "PR" , "RI" , "SC" , "SD" , "TN" , "TX" , "UT" , "VT" , "VA" , "WA" , "WV" , "WI" , "WY")

spattern <- data.frame(sname ,  states2)

# first iteration of states to convert state name to two letter abbreviation
for(s in spattern$sname){
  States <- agrep(s ,  US_States , ignore.case = TRUE, max.distance = 1)
  US_States[States] <- toupper(spattern$states2[spattern$sname==s])
}

#ST <- as.data.frame(US_States) %>% distinct()

# helper function to check if state pattern exists in the string
return_logic <- function(spattern, string){
  b <- FALSE
  w <- strsplit(string, "[, ;.]+") #%>% unlist()
  for (c in w){
    if(2 %in%str_length(c) && TRUE %in% grepl(spattern, c, ignore.case = TRUE)){
      b <- TRUE
    }
  }
  returnValue(b)
}
# second iteration of states to convert state name to two letter abbreviation
for (s in spattern$states2) {
  for (x in US_States) {
    if (return_logic(s,x) == TRUE){
      US_States[US_States==x] <- s
    }
  }
}

ST <- as.data.frame(US_States) %>% distinct() %>% arrange(US_States)
head(ST)
# Overwrite the result of states cleaned
file <- file %>% mutate(States = US_States)
```

### Cleaning Up Country Column


```{r}
# Clean up country names 
Countries <- file$Q4..COUNTRY %>% toupper() %>% str_trim()

# Pattern to match
pat <- c("USA","U.S.A", "U.S","United States","Costa Rica","America")
# Other countries
opat <- c("England", "Canada", "Mexico","Germany")

# Match pattern where several instance of misspelled United States of America
# Assign the Country to America
for (p in pat){
  matches <- agrep(p, Countries,ignore.case = TRUE, max.distance = 1)
  Countries[matches] <- "USA"
}
# Match pattern for other countries
for (p in opat){
  matches <- agrep(p, Countries,ignore.case = TRUE, max.distance = 1)
  Countries[matches] <- toupper(p)
}
# Some left the US State as the Country
for(s in spattern$sname){
  matches <- grep(s ,  Countries , ignore.case = TRUE)
  Countries[matches] <- "USA"
}

Co <- as.data.frame(Countries) %>% distinct() %>% arrange(Countries)

file <- file %>% mutate(Country = Countries)


```

### Extracting Valid Data
- Data where country is USA
- State is a valid USA State
- All the Candy with the ranking data

```{r}
# Remove special characters from the column names
names(file) <- gsub("Õ", "'", names(file))
names(file) <- gsub("\\."," ",names(file))
# Valid US data.
vd_US <- file %>% 
  mutate(notPerfect =  file %>% 
           select(starts_with("Q6")) %>%
           is.na %>% rowSums, .after = starts_with("Q5")) %>% 
  filter(Country == "USA" & States %in% spattern$states2 & notPerfect == 0) %>% 
  select(ID = 'Internal ID', 
         Gender = starts_with("Q2"), 
         Age = starts_with("Q3"),
         Country,
         States,
         starts_with("Q6"))  %>% 
 # rename_with(~gsub("^Q6   ","", .), starts_with("Q6"))
  pivot_longer(
    cols = starts_with("Q6"),
    names_to = "Candy_Catagory",
    values_to = "Rank",
    values_drop_na = TRUE
  )

# Extract the values for candy type
vd_US$Candy_Catagory <- gsub("Q6   ","",vd_US$Candy_Catagory)
# Convert Rank as integer
vd_US$Rank <- as.integer(vd_US$Rank)
glimpse(vd_US)
```


```{r}
# Candy Rank
candy_rank <- vd_US %>% 
  group_by(Candy_Catagory) %>% 
  summarize(Ranking = sum(Rank)) %>% 
  arrange(desc(Ranking))

# Plot the first 10 candy categories from rank
candy_rank %>% 
  head(10) %>%  
  ggplot(aes(y = Candy_Catagory, x = Ranking)) +
  geom_bar(stat = "identity") + 
  labs(
    y = "Candy Category",
    x = "Score",
    title = "Most Wanted Candy Category for Halloween 2017"
  )
```

