---
title: "DATA607_Assignment_9"
author: "Ying Fang Lee"
date: "2024-10-27"
output: html_document
---
## Assignment - Web APIs

### Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

We will need to use libraries httr, jsonlite, and dplyr to extract the data from API from JSON format to R Dataframe
```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(tidyverse)
```

Sign up for the NY Times developer apis and activate an API key for the API accessing. I choose the books API and from the API documentation get the base url and use it to build out API call. 
```{r}
# Define API key and endpoint
api_key <- "PR6ullqlzC8T72wCHIUWBbqkTQWAgRRe"
base_url <- "https://api.nytimes.com/svc/books/v3/lists/full-overview.json"
```

Get the response with the url constructed and check the response code
```{r}
# Construct the full URL with query parameters
url <- paste0(base_url, "?", "&api-key=", api_key) %>% print()

# Make the GET request
response <- GET(url)

# Check if the request was successful
if (status_code(response) != 200) {
  stop("API request failed!")
  }
```

Parse the response data into json format then convert it to dataframe.

```{r}
# Parse the JSON response into a list
json_data <- fromJSON(content(response, "text", encoding = "UTF-8"))

# Extract JSON to dataframe
book_lists <- json_data$results$lists %>% as.data.frame()
books <- json_data$results$lists$books
# Empty dataframe for books
books_df <- data.frame(counter = integer(),
                       title = character(), 
                       author = character(), 
                       rank = integer(), 
                       last_rank = integer(),
                       publisher = character(),
                       stringsAsFactors = FALSE)
# create an identifier for join later
book_lists <- book_lists %>% mutate(counter = 1:18)

# parsing the list of books
item <- 1
for (book in books) {
  dataf <- data.frame(
    counter = item,
    title = book$title,
    author = book$author,
    rank = book$rank,
    last_rank = book$rank_last_week,
    publisher = book$publisher,
    stringsAsFactors = FALSE
  )
  books_df <- rbind(books_df, dataf)
  item <- item+1
  }

# Create data frame for best sellers
best_sellers <- left_join(books_df, book_lists %>% select(counter, list_id,list_name,updated)) %>% select(-counter)

```
### Analysis

```{r}
# Which author was on the best seller list the most times
best_author <- best_sellers %>% 
  count(author) %>%
  arrange(desc(n)) %>%
  head(6)

best_author

best_sellers %>% 
  filter(author %in% best_author$author) %>% 
  ggplot(aes(x = author)) +
  geom_bar()
```


```{r}
# Which publisher produced the most best sellers
best_publisher <- best_sellers %>% 
  count(publisher) %>% 
  arrange(desc(n)) %>% 
  head(3)

best_publisher

best_sellers %>% 
  filter(publisher %in% best_publisher$publisher) %>% 
  ggplot(aes(x = publisher)) +
  geom_bar()
```


### Conslusion 
Freida McFadden	and	Kristin Hannah authored the most best sellers for different categories as of the list from December. Scholastic is the publisher that published the most best sellers. 

