---
title: "DATA607_Assignment_9"
author: "Ying Fang Lee"
date: "2024-10-27"
output: html_document
---
## Assignment - Web APIs

### Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

We will need to use libraries httr, jsonlite, and dplyr to extract the data from API from JSON format to R Dataframe
```{r}
library(httr)
library(jsonlite)
library(dplyr)
```

Sign up for the NY Times developer apis and activate an API key for the API accessing. I choose the books API and from the API documentation get the base url and use it to build out API call. 
```{r}
# Define API key and endpoint
api_key <- "PR6ullqlzC8T72wCHIUWBbqkTQWAgRRe"
base_url <- "https://api.nytimes.com/svc/books/v3/lists/names.json"
```

Get the response with the url constructed and check the response code
```{r}
# Construct the full URL with query parameters
url <- paste0(base_url, "?q=", "&api-key=", api_key) %>% print()

# Make the GET request
response <- GET(url)

# Check if the request was successful
if (status_code(response) != 200) {
  stop("API request failed!")
  }
```

Parse the response data into json format then convert it to dataframe.

```{r}
# Parse the JSON response into a list
json_data <- fromJSON(content(response, "text", encoding = "UTF-8"))

# Extract the articles from the JSON data
booknames <- json_data$results %>% as.data.frame()
head(booknames)
summary(booknames)
```

### Conclusion
There are 59 categories of NY Times best sellers. They are updated monthly or weekly where the oldest published book on the list starts from 2008. 