---
title: "DATA607_Assignment_4"
author: "Ying Fang Lee"
date: "9/29/2024"
output: html_document
---

### 1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

I have created a csv file with the information and uploaded to [Github](https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_4/Tifdying%20and%20Transforming%20Data.csv). I will be using this file to manipulate into tidy form.

```{r}
library(tidyverse)

# Read the file from GitHub and save to a data frame
airlines <- read.csv("https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_4/Tifdying%20and%20Transforming%20Data.csv")
print(airlines)
```


### 2) Tidy data
```{r}
# Taking care of empty spaces and set them to NA
airlines[airlines ==""] <- NA

# Looking at data using pivot_longer
airlines_long <- airlines %>% 
  fill(X, .direction = "down") %>%  #fill in the data from the previous row
  pivot_longer(
    cols = Los.Angeles:Seattle,
    names_to = "Location",
    values_to = "Count",
    values_drop_na = TRUE
  ) %>% 
  rename(Airline = X, Status = X.1) %>% 
  print()

# Looking at data using pivot_wider
airlines_wide <- airlines %>% 
  fill(X, .direction = "down") %>% 
  drop_na() %>% 
  pivot_wider(
    names_from = X.1,
    values_from = Los.Angeles:Seattle
  ) %>% 
  print()

```


