---
title: "DATA607_Assignment_4"
author: "Ying Fang Lee"
date: "9/29/2024"
output: html_document
---

### 1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

I have created a csv file with the information and uploaded to [Github](https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_4/Tifdying%20and%20Transforming%20Data.csv). I will be using this file to manipulate into tidy form.

```{r}
library(tidyverse)
library(janitor)

# Read the file from GitHub and save to a data frame
airlines <- read.csv("https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Assignment_4/Tifdying%20and%20Transforming%20Data.csv")
print(airlines)
```


### 2) Tidy data

To make the data tidy we would have to think about reformatting the table and present it with a way that each column is a variable, each row is an observation, and each cell is a value. First we have to think about the empty values in our table. I used excel to create my table and for the cells that have no value I tabbed to the next cell or I traverse the cells with the arrows. These cells were either recorded as empty cells "" or null values when converted to csv files. R has many build in functions that can take care of NA values, so the first step I did to clean up my data is to remove all empty cells and making them NA. Since my data did not have any other odd characters I did not do any more cleaning other than converting the empty cells to NA values. However, in other cases I think we would have to think about handling extra spaces, special characters, or corrupt data with regular expression. 

```{r}
# Taking care of empty spaces and set them to NA
airlines[airlines ==""] <- NA
```

Then I began organizing my data. The empty columns were renamed column X and X.1 after R load the data into data frame. I renamed my columns to meaningful names and correct formatting.

```{r}
# Using clean_names() part of the janitor package to replace the . in the names and remove space for the column name
airlines <- airlines %>% 
  rename(Airline = X, Status = X.1) %>% 
  clean_names()

# Using fill() part of dplyer to fill the data in the column with the value from before until a different value is encountered
airlines <- airlines %>% 
  fill(airline, .direction = "down")

print(airlines)
```
I do notice there is a row of NA values in my table but I know that R and the tidyverse library can handle this very easily. I tried looking at the data the "long" way and "wide" way with the pivot_longer and pivot_wider function.  
In pivot_longer, I condensed the columns and made a "location" column to include all the locations the flights arrive to and a "count" column to count the number of flights on time or delayed. Using the values_drop_na variable in pivot_longer functiion, I was able to drop the row with NA values. 
```{r}
# Looking at data using pivot_longer
airlines_long <- airlines %>% 
  pivot_longer(
    cols = los_angeles:seattle,
    names_to = "location",
    values_to = "count",
    values_drop_na = TRUE
  ) %>% 
  print()


```
In pivot_wider I dropped the row with NA values before I expand the columns with pivot_wider since I am trying to expand with the "status" variable against all the other location columns. I ended up with location columns of "on time" or "delayed"
```{r}
# Looking at data using pivot_wider
airlines_wide <- airlines %>% 
  drop_na() %>% # drop the rows that has na values in there
  pivot_wider(
    names_from = status,
    values_from = los_angeles:seattle
  ) %>% 
  print()
```

I think it is easier with pivot_longer to manipulate data later since I can think in columns vs thinking in rows. 

### Perform analysis to compare the arrival delays for the two airlines.
I tried two different ways of analyzing with the "longer" representation of data to compare the arrival delays of each airline. I think it would be best to represent as percentage of flights on time over all flights taken. First attempt I break the data into two tables and then join them to calculate the percentage. Second attempt I tried not breaking into other tables and used only group_by, mutate, and summarise to come up with final data. In my first attempt I can see all the calculated columns and gave me more insight but the result of both attempts are the same. 

Around 89 percent of AM WEST flights arrive on time and around 87 percent of Alaska flights arrive to the locations on time.

```{r}
# Break into different summary tables 
total_flights <- airlines_long %>% 
  group_by(airline) %>% 
  summarise(all_loc = sum(count)) #%>% 
total_status <- airlines_long %>% 
  group_by(airline, status) %>% 
  summarise(total_status = sum(count))
attempt_1 <- total_status %>%
  left_join(total_flights) %>% 
  mutate(percentage = total_status/all_loc*100) %>% 
  print()

# Second attempt without breaking into different tables
attempt_2 <- airlines_long %>% 
  group_by(airline) %>% 
  mutate(all_loc = sum(count),
         percentage = count/all_loc*100) %>% 
  group_by(airline, status, all_loc) %>% 
  summarise(total_per = sum(percentage)) %>% 
  print()

```

